---
title: "STAT510_Project"
author: "Nicholas Roy"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

```{r, echo=FALSE} 
pkgs = c("alr4","rmarkdown", "faraway", "datasets", "sos", "car", "plotly", "ggplot2", "tidyverse", "broom", "SemiPar", "RegularExpressions","GGally", "leaps", "usmap", "maps", "plotly", "devtools", "stargazer", "MASS")
new.packages = pkgs[!(pkgs %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(GGally)
library(ggplot2)
```

```{r}
##Working Directory
#setwd("C:\Users\mnhale\OneDrive - Urban Science\STATS\Spring 2020\STAT 510\covidanalysis\project")

#setwd("C:/Users/Nicho/OneDrive/2_Mathematics/School/courses/STAT510_intro_regression_analysis/project/github/covidanalysis/project")
```

#Data Preparation

##Step 1: Import Raw Datasets

```{r}

#casesanddeaths = read.csv(file = "csvs/us-counties.csv")
casesanddeaths = read.csv(url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"))
data_health = read.csv(file = "csvs/us-county-health-rankings-2020.csv")
SAHorder = read.csv(file = "csvs/state-stay-home.csv")
global_mobility = read.csv(file = "csvs/Global_Mobility_Report.csv")
#global_mobility = read.csv(url("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv?cachebust=a88b56a24e1a1e25"))

#***REMOVE ONCE NEW DATA IS WORKING***
data_mobility_old = read.csv(file = "csvs/us-mobility.csv")
data_mobility_old$county = stringr::str_remove(data_mobility$region, " County")
#***REMOVE ONCE NEW DATA IS WORKING***

```


##Step 2: Create US County Level Data
```{r}

#Clean
us_mobility = global_mobility

#Remove Non-US Countries
othercountries<-which(us_mobility$country_region_code!="US")
us_mobility<-us_mobility[-othercountries,]

#Remove Observations with Missing Country Data
nacountries<-which(is.na(us_mobility$country_region_code))
us_mobility<-us_mobility[-nacountries,]

#Remove non-county subregion_2 observations (i.e. cities, state level observations)

us_mobility = us_mobility[grepl("County", us_mobility$sub_region_2),]

```

##Step 3: Remove non SAH states

For this class we are limited to cross-sectional analysis. To address the inherent time-series nature of this data, we are going to perform a transformation that simplifies the data to a 4 week period and only looks at states that stayed at home. Because each state implemented the stay at home orders at different times, our observations are simplified to the change over the 4 week period immediately following the stay at home orders. According to the <a href="https://www.kff.org/other/slide/when-state-stay-at-home-orders-due-to-coronavirus-went-into-effect/"> Kaiser Family Foundation (KFF) </a>, Missouri and South Carolina were the last states to implement these orders on April 7th, 2020. We used the timeline from KFF to determine our observation period. Note according to <a href="https://www.usatoday.com/story/news/nation/2020/03/30/coronavirus-stay-home-shelter-in-place-orders-by-state/5092413002/"> USA Today </a> , we see that Wyoming, Utah, South Dakota, Oklahoma, North Dakota, Nebraska, Iowa, and Arkansas have not declared stay at home orders. Since these states have very few counties and cases, we will be excluding them from the mobility data prior to the analysis as they could be potential outliers. This only applies for the mobility variables as they are they are the only independent variables that are most likely to be affected by stay at home orders. The health variables are based on pre-pandemic data and thus will not suffer from this issue. For the dependent variable we look at the 4 week change in cases/population after the most day following the most recent statewide order in the nation (4/8/20). A separate analysis excluding these states will also be included.


```{r}
#Remove non SAH states

nonSAH_states = which(us_mobility$sub_region_1 == "Wyoming" | us_mobility$sub_region_1 == "Utah" | us_mobility$sub_region_1 == "South Dakota" | us_mobility$sub_region_1 == "Oklahoma" |us_mobility$sub_region_1 == "North Dakota" | us_mobility$sub_region_1 == "Nebraska" |us_mobility$sub_region_1 == "Iowa" | us_mobility$sub_region_1 == "Arkansas") #States w/ no stay at home order

us_mobility<-us_mobility[-nonSAH_states,]

```


##Step 4: Merge Mobility & Case Data
```{r}

#Clean Mobility Data to Merge w/ COVID Data
us_mobility$state = us_mobility$sub_region_1
us_mobility$county = stringr::str_remove(us_mobility$sub_region_2, " County")

CD_mobility = merge(x = casesanddeaths, y = us_mobility, all.x = TRUE, by = c("county", "state", "date"))

```

##Step 5: Adjust data based on "Stay at Home" orders

```{r}

SAHdates_withdata = merge(x = CD_mobility, y = SAHorder, all.x=TRUE, by.x = "state", by.y = "STATE")

#Get rid of pre stay at home order dates
earlydates = which(as.integer(as.Date(SAHdates_withdata$date)) < as.integer(as.Date("2020-03-18")))
SAHdates_withdata2 = SAHdates_withdata[-earlydates,]

#Focus on key dates
unimportant_dates = which(as.Date(SAHdates_withdata2$date) != as.Date(SAHdates_withdata2$EFF_DATE) & (as.Date(SAHdates_withdata2$date) - as.Date(SAHdates_withdata2$EFF_DATE) != 28)) #Only have dates for a state based on day of stay at home order and 3 weeks after.
impdates = SAHdates_withdata2[-unimportant_dates,]

pseudoSAH_dates =  which((impdates$date != "2020-04-08" & impdates$date != "2020-05-06") & (impdates$state == "Wyoming" | impdates$state == "Utah" | impdates$state == "South Dakota" | impdates$state == "Oklahoma" | impdates$state == "North Dakota" | impdates$state == "Nebraska" |impdates$state == "Iowa" | impdates$state == "Arkansas" |impdates$state == "Guam" | impdates$state == "Northern Mariana Islands" | impdates$state == "Puerto Rico" | impdates$state == "Virgin Islands")) #For locations where we don't have SAHOrder data

twodatesastate = impdates[-pseudoSAH_dates,]
```

##Step 6: Make variables differences over the past month to make cross sectional dataset from time-series.

```{r}
#Generating differences
sorteddata = twodatesastate[order(as.integer(twodatesastate$fips), as.Date(twodatesastate$date)),] #sort by fips and date
head(sorteddata)

require("data.table")
sorteddata <- data.table(sorteddata)
sorteddata[, case_change := c(NA,diff(cases)), by="fips"]
sorteddata[, retail_and_rec_pctpoint_change := c(NA,diff(retail_and_recreation_percent_change_from_baseline)), by="fips"]
sorteddata[, grocery_and_pharmacy_pctpoint_change := c(NA,diff(grocery_and_pharmacy_percent_change_from_baseline)), by="fips"]
sorteddata[, parks_pctpoint_change := c(NA,diff(parks_percent_change_from_baseline)), by="fips"]
sorteddata[, transit_stations_pctpoint_change := c(NA,diff(transit_stations_percent_change_from_baseline)), by="fips"]
sorteddata[, workplaces_pctpoint_change := c(NA,diff(workplaces_percent_change_from_baseline)), by="fips"]
sorteddata[, residential_pctpoint_change := c(NA,diff(residential_percent_change_from_baseline)), by="fips"]

missingcasedata = which(is.na(sorteddata$case_change))
almostfinalobs = sorteddata[-missingcasedata,]
missingfipsdata = which(is.na(almostfinalobs$fips))

finalobs = almostfinalobs[-missingfipsdata,]

```

##Step7: Add Health Variables

```{r}

allmerged = merge(x=finalobs, y = data_health, all.x = TRUE, by = "fips")
allmerged$state = allmerged$state.x

```


##Step8: Make datasets with variables of interest.
```{r}

## Full Linear Models for Cases

alldata = subset(allmerged, select=c("fips","state","case_change","percent_low_birthweight","percent_smokers","primary_care_physicians_rate","percent_vaccinated","num_households","average_daily_pm2_5","life_expectancy","child_mortality_rate","infant_mortality_rate","percent_food_insecure","median_household_income","average_traffic_volume_per_meter_of_major_roadways","population_2","percent_65_and_over","percent_rural","percent_uninsured","workplaces_pctpoint_change", "retail_and_rec_pctpoint_change","grocery_and_pharmacy_pctpoint_change","parks_pctpoint_change","transit_stations_pctpoint_change","residential_pctpoint_change", "average_traffic_volume_per_meter_of_major_roadways"))

healthsubset = subset(allmerged,select=c("fips","state","case_change","percent_low_birthweight","percent_smokers","primary_care_physicians_rate","percent_vaccinated","num_households","average_daily_pm2_5","life_expectancy","child_mortality_rate","infant_mortality_rate","percent_food_insecure","median_household_income","population_2","percent_65_and_over","percent_rural","percent_uninsured"))

#For checking which mobility variable will be og greatest interest
# summary(alldata$retail_and_rec_pctpoint_change)
# summary(alldata$grocery_and_pharmacy_pctpoint_change) 
# summary(alldata$parks_pctpoint_change)
# summary(alldata$transit_stations_pctpoint_change)
# summary(alldata$workplaces_pctpoint_change)
# summary(alldata$residential_pctpoint_change)

```



## Step 9: Check Correlation between variables and also linearity between independent variables and the dependent variable
```{r}
#based on the output here, we will probably need to log transform casesmonth
ggpairs(healthsubset, columns = c(3, 4:9))
ggpairs(healthsubset, columns = c(3, 10:14))
ggpairs(healthsubset, columns = c(3, 14:18))
#head(cov19_project_nm)
```

We can see that there is non-normality of the case change per capita variable. We will likely log transform this variable. Immdiately we see some variables have unequal variance, but we will perform variable selection before making any transformations.

As far as variables that are collinear, we make the following observations:
We see that life expectancy is collinear with many of the variables. This is to be expected since it is a more general term that is determined by the other variables.This wouldn't be useful as an interaction term, so we will look at the scatterplot matrix of variables correlated with life expectancy.


## Step 10: Look at correlation with life expectancy.


```{r}
ggpairs(healthsubset, columns = c(10, 4,5,11,12,13,14))
#head(cov19_project_nm)
```

Based off this information we can see that median household income might be the interaction with these variables. Life expectancy could be collinear with percent smokers and percent food insecure, but applying our variable selection criteria should remove the variable with the least amount of explanatory power. We should look for interactions with these variables after the variable selection.If interactions from here are not significant later in the selection process than we will remove collinear terms based on their signifigance.

## Step 11: Use Box-cox method to determine best transformation for the dependent variable. The variable tempcases is simply to be positive to prevent errors for the box cox transformation

```{r}
library(car)

summary(healthsubset$case_change)
hist(healthsubset$case_change)

healthsubset$tempcases = healthsubset$case_change + 1.5 #Only to remove negative issue from boxcox and for future log transform

boxcoxtestreg = lm(data = na.omit(healthsubset), tempcases ~  percent_low_birthweight +percent_smokers +primary_care_physicians_rate + num_households +average_daily_pm2_5 +life_expectancy +infant_mortality_rate +median_household_income +percent_65_and_over + percent_rural )

bc = boxCox(boxcoxtestreg)

lambda.opt = bc$x[which.max(bc$y)]
lambda.opt

#Confirms that log transform is the best for cases


healthsubset$logcasechange = log(healthsubset$tempcases) #usingtemp cases to remove places that reduced cases

```

From the box-cox method, we can see that a log transformation is most appropriate.

## Step 12: We can now do variable selection using the AIC variable selection criteria.

```{r}

mod0.lower_cases = lm(data = na.omit(healthsubset), logcasechange ~ 1)
mod0.upper_cases = lm(data = na.omit(healthsubset), logcasechange ~  percent_low_birthweight + percent_smokers +primary_care_physicians_rate +percent_vaccinated + num_households +average_daily_pm2_5 +life_expectancy +child_mortality_rate +infant_mortality_rate +percent_food_insecure +median_household_income +percent_65_and_over +percent_rural +percent_uninsured)
step(mod0.lower_cases, scope = list(lower = mod0.lower_cases, upper = mod0.upper_cases))

```

## Step 13: We now have the base model for what variables we will be using. Now we must check the 4 assumptions of Multiple Linear Regression: Linearity, Independence, Normality, and Equal Variance


```{r}

lm1 = lm(formula = logcasechange ~ percent_rural + num_households + 
    median_household_income + percent_low_birthweight + percent_vaccinated + 
    infant_mortality_rate + percent_uninsured + life_expectancy + 
    average_daily_pm2_5 + child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(residuals(lm1), fitted(lm1))

shapiro.test(residuals(lm1)) #Doesn't pass normality
hist(residuals(lm1)) #Looks relatively normal
plot(lm1, which = 2) #Looks relatively normal


```
We appear to have decent equal variance, linearity, and assume independence. However our model still does not satisfy the normality condition. To resolve this issue we will look to see if each variable satisfies equal variance conditions and then recheck normality.

```{r}

plot(y = residuals(lm1), x = na.omit(healthsubset)$percent_rural) # relatively equal variance. appears there is one particular outlier

plot(y = residuals(lm1), x= na.omit(healthsubset)$num_households) # clear funnelling patter that could use a log transform.

healthsubset$loghouseholds = log(healthsubset$num_households)

lm2 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    median_household_income + percent_low_birthweight + percent_vaccinated + 
    infant_mortality_rate + percent_uninsured + life_expectancy + 
    average_daily_pm2_5 + child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y = residuals(lm2), x=na.omit(healthsubset)$loghouseholds) #Still some minor funnelling, but less severe than in the previous graph

plot(y = residuals(lm2), x= na.omit(healthsubset)$median_household_income) #some minor funnelling, log transform might help


healthsubset$logmedincome = log(healthsubset$median_household_income)

lm3 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + percent_low_birthweight + percent_vaccinated + 
    infant_mortality_rate + percent_uninsured + life_expectancy + 
    average_daily_pm2_5 + child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y = residuals(lm3), x=na.omit(healthsubset)$logmedincome) #Equal Variance Acheived

plot(y = residuals(lm3), x=na.omit(healthsubset)$percent_low_birthweight) #some minor funnelling, log transform might help


healthsubset$log_percent_low_birthweight = log(healthsubset$percent_low_birthweight)

lm4 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + log_percent_low_birthweight + percent_vaccinated + 
    infant_mortality_rate + percent_uninsured + life_expectancy + 
    average_daily_pm2_5 + child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y = residuals(lm4), x=na.omit(healthsubset)$log_percent_low_birthweight) #Equal Variance Acheived

plot(y = residuals(lm4), x=na.omit(healthsubset)$percent_vaccinated) #Seems like there is equal variance
plot(y = residuals(lm4), x=na.omit(healthsubset)$infant_mortality_rate) #Needs log transform

healthsubset$log_infant_mortality_rate = log(healthsubset$infant_mortality_rate)

lm5 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + log_percent_low_birthweight + percent_vaccinated + 
    log_infant_mortality_rate + percent_uninsured + life_expectancy + 
    average_daily_pm2_5 + child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y = residuals(lm5), x=na.omit(healthsubset)$log_infant_mortality_rate) #equal variance acheived

plot(y = residuals(lm5), x=na.omit(healthsubset)$percent_uninsured) #Minor Funnelling, could use log transform

healthsubset$log_percent_uninsured = log(healthsubset$percent_uninsured)

lm6 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + log_percent_low_birthweight + percent_vaccinated + 
    log_infant_mortality_rate + log_percent_uninsured + life_expectancy + 
    average_daily_pm2_5 + child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y=residuals(lm6), x=na.omit(healthsubset)$log_percent_uninsured) #Improved funnelling pattern to reach equal variance

plot(y=residuals(lm6), x=na.omit(healthsubset)$life_expectancy) #Looks like equal variance is acheived

plot(y=residuals(lm6), x=na.omit(healthsubset)$average_daily_pm2_5)#Funnelling pattern

healthsubset$log_average_daily_pm2_5 = log(healthsubset$average_daily_pm2_5) 

lm7 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + log_percent_low_birthweight + percent_vaccinated + 
    log_infant_mortality_rate + log_percent_uninsured + life_expectancy + 
    log_average_daily_pm2_5 + child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y=residuals(lm7), x=na.omit(healthsubset)$log_average_daily_pm2_5) #Improved funnelling pattern to reach near equal variance

plot(y=residuals(lm7), x=na.omit(healthsubset)$child_mortality_rate) #Needs log transformation


healthsubset$log_child_mortality_rate = log(healthsubset$child_mortality_rate)

lm8 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + log_percent_low_birthweight + percent_vaccinated + 
    log_infant_mortality_rate + log_percent_uninsured + life_expectancy + 
    log_average_daily_pm2_5 + log_child_mortality_rate + percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y=residuals(lm8), x=na.omit(healthsubset)$log_child_mortality_rate) #Fixed funnelling and achieved equal variance

plot(y=residuals(lm8), x=na.omit(healthsubset)$percent_food_insecure) #Funnelling pattern

healthsubset$log_percent_food_insecure = log(healthsubset$percent_food_insecure) 

lm9 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + log_percent_low_birthweight + percent_vaccinated + 
    log_infant_mortality_rate + log_percent_uninsured + life_expectancy + 
    log_average_daily_pm2_5 + log_child_mortality_rate + log_percent_food_insecure, 
    data = na.omit(healthsubset))

plot(y=residuals(lm9), x=na.omit(healthsubset)$log_percent_food_insecure) #Fixed funnelling and achieved equal variance

plot(lm9,which = 2)
hist(residuals(lm9))
plot(residuals(lm9), fitted(lm9))
shapiro.test(residuals(lm9))

```
We see that we satisfy equal variance and linearity conditions, but we have not acheived normality. We will check for if any variables previously excluded should now be included.


## Step 14: Perform variable selection test again but for all newly transformed variables

```{r}

lm9.upper_cases = lm(data = na.omit(healthsubset),logcasechange ~ percent_rural + loghouseholds + logmedincome + log_percent_low_birthweight + percent_vaccinated + log_infant_mortality_rate + log_percent_uninsured + life_expectancy + log_average_daily_pm2_5 + log_child_mortality_rate + log_percent_food_insecure + log_percent_uninsured + percent_smokers + primary_care_physicians_rate + percent_65_and_over)
step(lm9, scope = list(lower = lm9, upper = lm9.upper_cases))

```

Now we should include percent_smokers.

```{r}

lm10 = lm(formula = logcasechange ~ percent_rural + loghouseholds + 
    logmedincome + log_percent_low_birthweight + percent_vaccinated + 
    log_infant_mortality_rate + log_percent_uninsured + life_expectancy + 
    log_average_daily_pm2_5 + log_child_mortality_rate + log_percent_food_insecure + 
    percent_smokers, data = na.omit(healthsubset))

plot(y = residuals(lm10), x = na.omit(healthsubset)$percent_smokers) #It appears there is equal variance with a few outliers.

shapiro.test(residuals(lm10)) #Still have not solved the normality problems

```

Now we will test interaction terms with variable selection.


#Step 15: Use step function with interaction terms with log median income.


```{r, echo=FALSE}

library(MASS)

help(stepAIC)
lm10interactions = stepAIC(lm10, scope=list(upper= logcasechange ~ percent_rural + loghouseholds + logmedincome + log_percent_low_birthweight + percent_vaccinated + log_infant_mortality_rate + log_percent_uninsured + life_expectancy + log_average_daily_pm2_5 + log_child_mortality_rate + log_percent_food_insecure + percent_smokers + loghouseholds:percent_rural + loghouseholds:logmedincome + loghouseholds:log_percent_low_birthweight + loghouseholds:percent_vaccinated + loghouseholds:log_infant_mortality_rate + loghouseholds:log_percent_uninsured + loghouseholds:life_expectancy + loghouseholds:log_average_daily_pm2_5 + loghouseholds:log_child_mortality_rate + loghouseholds:log_percent_food_insecure + loghouseholds:percent_smokers, lower= ~1))

```


## Step 16: Check normality with new terms considered.
```{r}

summary(lm10interactions)
shapiro.test(residuals(lm10interactions))
plot(lm10interactions, which = 2)
hist(residuals(lm10interactions))

```
We still did not achieve normality, so we must remove outliers.
## Step 17: Remove outliers

```{r}
cooks = cooks.distance(lm10interactions)
n = nrow(na.omit(healthsubset))

plot(lm10interactions, which = 4)
abline(h = 1, lty = 2)

highLeverage = cooks.distance(lm10interactions)> (4/n)
highLeverage 

healthsubset.new = na.omit(healthsubset)[!highLeverage,]

lm10interactions.new = lm(logcasechange ~ percent_rural + loghouseholds + logmedincome + log_percent_low_birthweight + percent_vaccinated + log_percent_uninsured + life_expectancy + log_average_daily_pm2_5 + log_child_mortality_rate + log_percent_food_insecure + percent_smokers + loghouseholds:log_child_mortality_rate +  loghouseholds:log_percent_uninsured + loghouseholds:log_percent_low_birthweight +     loghouseholds:log_average_daily_pm2_5 + loghouseholds:life_expectancy +     percent_rural:loghouseholds + loghouseholds:log_percent_food_insecure, data = healthsubset.new)
lm10interactions.lower = lm(logcasechange ~ 1, data = healthsubset.new)

shapiro.test(residuals(lm10interactions.new))
summary(lm10interactions.new)
summary(lm10interactions)
```


## Step 18: General F-test for mobility variables

```{r}
anova(lm10interactions.lower, lm10interactions.new)

```

## Step 19: Output tables

```{r}



```


#Ctrl + Shift + C to uncomment code below.



<!-- ```{r} -->

<!-- #normality issue -->
<!-- #remove potential outliers (no time to use influential point methods) -->


<!-- plot(lm8, which = 2) -->
<!-- plot(residuals(lm8), fitted(lm8)) -->

<!-- cov19_project_nm_nooutliers = na.omit(cov19_project_nm) -->
<!-- cov19_project_nm_nooutliers$lm8resids = na.omit(residuals(lm8)) -->

<!-- outliers = boxplot(cov19_project_nm_nooutliers$lm8resids, plot=FALSE)$out -->

<!-- lm8resids = append(outliers, 5:925, after = length(outliers) - 1) -->

<!-- outliermarker = rep(1,length(cov19_project_nm_nooutliers)) -->
<!-- outliermat = as.data.frame(cbind(lm8resids, outliermarker)) -->

<!-- cov19_project_nm_nooutliers = merge(x = cov19_project_nm_nooutliers, y = outliermat, by = "lm8resids", all.x = TRUE) -->

<!-- cov19_project_nm_nooutliers<- cov19_project_nm_nooutliers[-which(cov19_project_nm_nooutliers$outliermarker == 1),] -->


<!-- lm9 = lm(data = cov19_project_nm_nooutliers, logcasesmonth ~  log_percent_low_birthweight + percent_smokers +log_primary_care_physicians_rate +average_daily_pm2_5 +life_expectancy +log_infant_mortality_rate +log_median_household_income + log_percent_65_and_over +majorityruralfactor) -->

<!-- plot(lm9, which = 2) -->

<!-- shapiro.test(residuals(lm9)) -->
<!-- hist(residuals(lm9)) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- library(usmap) -->
<!-- ggplot(data = cov19_project_nm, aes(log_percent_65_and_over, logcasesmonth, col = factor(state)))+ -->
<!--   geom_point()+ -->
<!--   labs(title = "Scatterplot Log(% over 65) vs. Log(Cases Month)", -->
<!--   x = "Log(% over 65)", -->
<!--   y= "Log(Cases Month)") -->

<!-- states = cov19_project_nm$STATE -->


<!-- plot_usmap(regions = c("states"), data = cov19_project_nm, -->
<!--            values = "casesmonth", color = "light gray", labels = FALSE, -->
<!--            label_color = "black") + scale_fill_continuous(low = "pink", high = "red", na.value = " gray", name ="", label = scales::comma) + -->
<!--   theme(legend.key.width = unit(2, "line"), legend.key.height = unit(3, "line"), legend.position = "right", legend.title = element_text(size=16)) -->

<!-- help(plot_usmap) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- #prediction interval for counties with 65+ -->
<!-- fit2 = lm(logcasesmonth ~ log_percent_low_birthweight + percent_smokers +  -->
<!--     log_primary_care_physicians_rate + average_daily_pm2_5 +  -->
<!--     life_expectancy + log_infant_mortality_rate + log_median_household_income +  -->
<!--     log_percent_65_and_over + majorityruralfactor + log_median_household_income:log_percent_65_and_over +  -->
<!--     log_median_household_income:majorityruralfactor + log_percent_low_birthweight:log_median_household_income +  -->
<!--     percent_smokers:log_median_household_income + life_expectancy:log_median_household_income,  -->
<!--     data = na.omit(cov19_project_nm)) -->
<!-- summary(fit1) -->
<!-- summary(fit2) -->

<!-- #looks like relatively equal variance on the right side -->
<!-- x10 = na.omit(cov19_project_nm)$log_median_household_income*na.omit(cov19_project_nm)$majorityruralfactor -->
<!-- plot(x = x10, y= na.omit(cov19_project_nm)$logcasesmonth,cex.axis = .6, xlab = "", ylab = "") -->
<!-- title(xlab = "log(Med HH Income)*majority rural", line = 2, cex.lab = .7) -->
<!-- title(ylab = "log(cases month)", line=2, cex.lab = .7) -->
<!-- title(main = "log(Med HH Income)*majority rural vs. log(cases month)",line = 0.5, cex.lab = .5 ) -->


<!-- x11 = na.omit(cov19_project_nm)$log_median_household_income*na.omit(cov19_project_nm)$log_percent_low_birthweight -->
<!-- plot(x = x11, y= na.omit(cov19_project_nm)$logcasesmonth,cex.axis = .6, xlab = "", ylab = "") -->
<!-- title(xlab = "log(Med HH Income)*log(% low birth weight)", line = 2, cex.lab = .7) -->
<!-- title(ylab = "log(cases month)", line=2, cex.lab = .7) -->
<!-- title(main = "log(Med HH Income)*log(% low birth weight) vs. log(cases month)",line = 0.5, cex.lab = .5 ) -->


<!-- x12 = na.omit(cov19_project_nm)$log_median_household_income*na.omit(cov19_project_nm)$percent_smokers -->
<!-- plot(x = x12, y= na.omit(cov19_project_nm)$logcasesmonth,cex.axis = .6, xlab = "", ylab = "") -->
<!-- title(xlab = "log(Med HH Income)*percent smokers", line = 2, cex.lab = .7) -->
<!-- title(ylab = "log(cases month)", line=2, cex.lab = .7) -->
<!-- title(main = "log(Med HH Income)*percent_smokers vs. log(cases month)",line = 0.5, cex.lab = .5 ) -->

<!-- x13 = na.omit(cov19_project_nm)$log_median_household_income*na.omit(cov19_project_nm)$life_expectancy -->
<!-- plot(x = x13, y= na.omit(cov19_project_nm)$logcasesmonth,cex.axis = .6, xlab = "", ylab = "") -->
<!-- title(xlab = "log(Med HH Income)*life_expectancy", line = 2, cex.lab = .7) -->
<!-- title(ylab = "log(cases month)", line=2, cex.lab = .7) -->
<!-- title(main = "log(Med HH Income)*life_expectancy vs. log(cases month)",line = 0.5, cex.lab = .5 ) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- #mobility -->
<!-- cov19_project$logcasesmonth = cov19_project_nm$logcasesmonth -->
<!-- cov19_project$log_percent_low_birthweight = log(cov19_project$percent_low_birthweight) -->
<!-- cov19_project$log_primary_care_physicians_rate = log(cov19_project$primary_care_physicians_rate) -->
<!-- cov19_project$log_infant_mortality_rate = log(cov19_project$infant_mortality_rate) -->
<!-- cov19_project$log_median_household_income = log(cov19_project$median_household_income) -->
<!-- cov19_project$log_percent_65_and_over = log(cov19_project$percent_65_and_over) -->
<!-- cov19_project$majorityruralfactor = ifelse(cov19_project_nm$percent_rural > .5, 1, 0) -->


<!-- lm8 = lm(data = na.omit(cov19_project_nm), logcasesmonth ~  log_percent_low_birthweight + percent_smokers +log_primary_care_physicians_rate + average_daily_pm2_5 + life_expectancy + log_infant_mortality_rate +log_median_household_income + log_percent_65_and_over + majorityruralfactor) -->

<!-- fit2 = lm(logcasesmonth ~ log_percent_low_birthweight + percent_smokers +  -->
<!--     log_primary_care_physicians_rate + average_daily_pm2_5 +  -->
<!--     life_expectancy + log_infant_mortality_rate + log_median_household_income +  -->
<!--     log_percent_65_and_over + majorityruralfactor + log_median_household_income:log_percent_65_and_over +  -->
<!--     log_median_household_income:majorityruralfactor + log_percent_low_birthweight:log_median_household_income +  -->
<!--     percent_smokers:log_median_household_income + life_expectancy:log_median_household_income,  -->
<!--     data = na.omit(cov19_project_nm)) -->

<!-- fit2_alt = lm(logcasesmonth ~ log_percent_low_birthweight + percent_smokers +  -->
<!--     log_primary_care_physicians_rate + average_daily_pm2_5 +  -->
<!--     life_expectancy + log_infant_mortality_rate + log_median_household_income +  -->
<!--     log_percent_65_and_over + majorityruralfactor + log_median_household_income:log_percent_65_and_over +  -->
<!--     log_median_household_income:majorityruralfactor + log_percent_low_birthweight:log_median_household_income +  -->
<!--     percent_smokers:log_median_household_income + life_expectancy:log_median_household_income,  -->
<!--     data = na.omit(cov19_project)) -->

<!-- fit2mobility = lm(data = na.omit(cov19_project), logcasesmonth ~ log_percent_low_birthweight + percent_smokers +  -->
<!--     log_primary_care_physicians_rate + average_daily_pm2_5 +  -->
<!--     life_expectancy + log_infant_mortality_rate + log_median_household_income +  -->
<!--     log_percent_65_and_over + majorityruralfactor + log_median_household_income:log_percent_65_and_over +  -->
<!--     log_median_household_income:majorityruralfactor + log_percent_low_birthweight:log_median_household_income +  -->
<!--     percent_smokers:log_median_household_income + life_expectancy:log_median_household_income + retail_recreation + grocery_pharmacy +parks + transit_stations + workplaces + residential) -->

<!-- shapiro.test(residuals(fit2)) -->
<!-- shapiro.test(residuals(fit2mobility)) -->

<!-- plot(fit2, which = 2) -->
<!-- hist(residuals(fit2), main = "Histogram of Model Residuals", xlab = "Standardized Residuals") -->

<!-- anova(fit2_alt, fit2mobility) -->
<!-- summary(lm8) -->
<!-- summary(fit2) -->
<!-- summary(fit2mobility) -->
<!-- library(stargazer) -->
<!-- help(stargazer) -->
<!-- table = stargazer(lm8, fit2, fit2mobility, title = "Results", align = TRUE, type ="html") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- plot(y=residuals(fit2), x= x10) #fixes unequal variance -->

<!-- p.i. = predict(fit2, interval = 'confidence', level = .95) -->
<!-- confint(fit2, "log_percent_65_and_over",level = .95) -->
<!-- summary(fit2) -->

<!-- mean(cov19_project$deathmonth) -->
<!-- ``` -->

